Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-10-20 doc/Paper/main.tex
Small changes to magnitude cuts section.

In Fig.~\ref{fig:HOD-redshift}, we additionally show the large-scale galaxy bias as a function of redshift derived from our best-fit HOD model. As can be seen, the large-scale bias scales approximately as $b(z) \sim \sfrac{1}{D(z)}$, where $D(z)$ denotes the linear growth factor. This behavior is consistent with e.g. studies of galaxies in the DEEP2 survey, which found constant clustering strengths as a function of redshift (and thus $b(z) \sim \sfrac{1}{D(z)}$) for magnitude-limited samples \cite{Coil:2004}. \todo{How about saying that a more thorough analysis of the color-dependence of clustering properties would be needed to fully characterize this effect and that we leave this for future work?}

\begin{figure}
\begin{center}
\subfigure{\includegraphics[width=0.32\textwidth]{figures/M_halo_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}} 
\subfigure{\includegraphics[width=0.32\textwidth]{figures/bias-redshift_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}}
\subfigure{\includegraphics[width=0.32\textwidth]{figures/bias-growth_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}} 
\caption{Redshift dependence of the mean halo mass $\bar{M}_{\mathrm{halo}}$ and the large-scale galaxy bias $b_{\mathrm{HOD}}$ as derived from our best-fit HOD model. The third panel illustrates the empirical scaling $b(z) \sim \sfrac{1}{D(z)}$ derived from the galaxy sample used in this analysis.}
\label{fig:HOD-redshift}
\end{center}
\end{figure}

@@

To this end, we split our fiducial galaxy sample into four brighter subsamples with limiting magnitudes $i_{\mathrm{corr}} < 20.5$, $i_{\mathrm{corr}} < 21.5$, $i_{\mathrm{corr}} < 22.5$ and $i_{\mathrm{corr}} < 23.5$, respectively. We then subdivide each of these samples into four redshift bins of approximately equal galaxy number and use our pipeline to compute constraints on HOD parameters for our fiducial analysis variant. We find all subsamples to be well-fit by our fiducial model described in Sec.~\ref{ssec:methods.theory}. In fact, the reduced $\chi^{2}$s for some subsamples are quite low and thus lead to correspondingly high $p$-values, which seem to indicate the presence of overfitting and/or covariance overestimation in our brighter magnitude samples. However, for all subsamples, the distribution of the fit residuals is consistent with a Gaussian and we do not find any significant evidence that our covariance is overestimated by comparing the consistency of power spectra across different HSC fields. We therefore use the results of these HOD fits to derive constraints on the large-scale galaxy bias as a function of redshift for each sample. We determine the bias from the mean of the posterior distribution, as we find it to yield more stable constraints than the best-fit value. The results for the bias evaluated at the effective redshift of each tomographic bin are shown in Fig.~\ref{fig:magnitude-cuts-bias}, alongside the results for our fiducial sample. As can be seen, we find the linear bias to increase for decreasing limiting magnitude, as expected. Furthermore, the redshift-dependence of the bias can be roughly approximated as inversely proportional to the growth factor, $b(z) \sim \sfrac{1}{D(z)}$, for all samples. In order to obtain a fitting function for the large-scale galaxy bias as a function of redshift and limiting magnitude, we assume it to be separable in $z$ and $m_{\mathrm{lim}}$ and make the following Ansatz:
\begin{equation}
b(z, m_{\mathrm{lim}}) = \bar{b}(m_{\mathrm{lim}}) D(z)^{\alpha}.
\label{eq:bias-fit-func}
\end{equation}  
We determine the values of $\alpha$ and $\bar{b}(m_{\mathrm{lim}})$ in a three-step weighted, linear least squares fit: we first fit Eq.~\ref{eq:bias-fit-func} to all five samples separately and determine our fiducial $\alpha$ as the median of these values\footnote{We note that we find slightly better results using the median, which is the reason for preferring it over the mean.}. We then fit Eq.~\ref{eq:bias-fit-func} to all samples again, obtaining constraints on $\bar{b}(m_{\mathrm{lim}})$ for $\alpha$ fixed to our fiducial value. In a last step, we fit a linear function to $\bar{b}(m_{\mathrm{lim}})$, finally yielding
\begin{equation}
\begin{aligned}
\alpha &= -1.32, \\
\bar{b}(m_{\mathrm{lim}}) &= b_{1}m_{\mathrm{lim}} + b_{0},
\label{eq:bias-fit-func-result}
\end{aligned}  
\end{equation}
where $b_{1} = -0.0627$ and $b_{0} = 2.35$. \as{Do we want to rewrite this as $\bar{b} = b_0 + b_1 (m-24)$?} The results from this fitting function are shown as solid lines in Fig.~\ref{fig:magnitude-cuts-bias}, alongside the values obtained from the HOD fit.

This analysis yields a few interesting results. First we find that the approximation $b(z)\propto D(z)^{-1}$ (i.e $\alpha=-1$ is only approximately valid. In fact, there is no physical reason for it, because the cut is on the observed magnitude, so at the same observed magnitude cut, the low redshift galaxies are \emph{different population} of galaxies. We find that the bias increases somewhat faster than the standard empirical assumption of $D(z)^{-1}$.

The most common use of this results is for Fisher matrix forecasts for constraining power of magnitude-limited samples. \as{Do we want to show our result for LSST vs SRD assumption of $0.95/D(z)$?}

An alternative use is galaxy evolution studies, where this results provides a convenient cross-check formula. Our full parameter chains for different magnitudes are available upon request. \as{Is this the case?} \an{That would be totally fine with me.}

>>>>

In Fig.~\ref{fig:HOD-redshift}, we additionally show the large-scale galaxy bias as a function of redshift derived from our best-fit HOD model. As can be seen, the large-scale bias scales approximately as $b(z) \propto \sfrac{1}{D(z)}$, where $D(z)$ denotes the linear growth factor. This behavior is consistent with e.g. studies of galaxies in the DEEP2 survey, which found constant clustering strengths as a function of redshift (and thus $b(z) \propto \sfrac{1}{D(z)}$) for magnitude-limited samples \cite{Coil:2004}. \todo{How about saying that a more thorough analysis of the color-dependence of clustering properties would be needed to fully characterize this effect and that we leave this for future work?}

\begin{figure}
\begin{center}
\subfigure{\includegraphics[width=0.32\textwidth]{figures/M_halo_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}} 
\subfigure{\includegraphics[width=0.32\textwidth]{figures/bias-redshift_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}}
\subfigure{\includegraphics[width=0.32\textwidth]{figures/bias-growth_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}} 
\caption{Redshift dependence of the mean halo mass $\bar{M}_{\mathrm{halo}}$ and the large-scale galaxy bias $b_{\mathrm{HOD}}$ as derived from our best-fit HOD model. The third panel illustrates the empirical scaling $b(z) \propto \sfrac{1}{D(z)}$ derived from the galaxy sample used in this analysis.}
\label{fig:HOD-redshift}
\end{center}
\end{figure}

@@

To this end, we split our fiducial galaxy sample into four brighter subsamples with limiting magnitudes $i_{\mathrm{corr}} < 20.5$, $i_{\mathrm{corr}} < 21.5$, $i_{\mathrm{corr}} < 22.5$ and $i_{\mathrm{corr}} < 23.5$, respectively. We then subdivide each of these samples into four redshift bins of approximately equal galaxy number and use our pipeline to compute constraints on HOD parameters for our fiducial analysis variant. We find all subsamples to be well-fit by our fiducial model described in Sec.~\ref{ssec:methods.theory}. In fact, the reduced $\chi^{2}$s for some subsamples are quite low and thus lead to correspondingly high $p$-values, which seem to indicate the presence of overfitting and/or covariance overestimation in our brighter magnitude samples. However, for all subsamples, the distribution of the fit residuals is consistent with a Gaussian and we do not find any significant evidence that our covariance is overestimated by comparing the consistency of power spectra across different HSC fields. We therefore use the results of these HOD fits to derive constraints on the large-scale galaxy bias as a function of redshift for each sample. We determine the bias from the mean of the posterior distribution, as we find it to yield more stable constraints than the best-fit value. The results for the bias evaluated at the effective redshift of each tomographic bin are shown in Fig.~\ref{fig:magnitude-cuts-bias}, alongside the results for our fiducial sample. As can be seen, we find the linear bias to increase for decreasing limiting magnitude, as expected. Furthermore, the redshift-dependence of the bias can be roughly approximated as inversely proportional to the growth factor, $b(z) \propto \sfrac{1}{D(z)}$, for all samples, as can be seen from the middle panel of Fig.~\ref{fig:magnitude-cuts-bias}. Motivated by these results, we make the following Ansatz to derive a fitting function for the large-scale galaxy bias as a function of redshift and limiting magnitude:
\begin{equation}
b(z, m_{\mathrm{lim}}) = \bar{b}(m_{\mathrm{lim}}) D(z)^{\alpha},
\label{eq:bias-fit-func}
\end{equation}  
where we have assumed the bias to be separable in $z$ and $m_{\mathrm{lim}}$ and the parameter $\alpha$ accounts for deviations to the simple inverse growth function proportionality.
We determine the values of $\alpha$ and $\bar{b}(m_{\mathrm{lim}})$ in a three-step weighted, linear least squares fit: we first fit Eq.~\ref{eq:bias-fit-func} to all five samples separately and determine our fiducial $\alpha$ as the median of these values\footnote{We note that we find slightly better results using the median, which is the reason for preferring it over the mean.}. We then fit Eq.~\ref{eq:bias-fit-func} to all samples again, obtaining constraints on $\bar{b}(m_{\mathrm{lim}})$ for $\alpha$ fixed to our fiducial value. In a last step, we fit a linear function to $\bar{b}(m_{\mathrm{lim}})$, finally yielding
\begin{equation}
\begin{aligned}
\alpha &= -1.32, \\
\bar{b}(m_{\mathrm{lim}}) &= b_{1}m_{\mathrm{lim}} + b_{0},
\label{eq:bias-fit-func-result}
\end{aligned}  
\end{equation}
where $b_{1} = -0.0627$ and $b_{0} = 2.35$. \as{Do we want to rewrite this as $\bar{b} = b_0 + b_1 (m-24)$?} The results from this fitting function are shown as solid lines in Fig.~\ref{fig:magnitude-cuts-bias}, alongside the values obtained from the HOD fit.

This analysis yields a few interesting results. First, we find that the approximation $b(z) \propto D(z)^{-1}$ (i.e $\alpha=-1$) is only approximately valid. In fact, there is no physical reason for it, because the cut is on the observed magnitude, so at the same observed magnitude cut, the low redshift galaxies are \emph{different population} of galaxies. We find that the bias increases somewhat faster than the standard empirical assumption of $D(z)^{-1}$.

We anticipate that these results will mostly be useful for forecasting the  constraining power of magnitude-limited samples in Fisher matrix forecasts. \as{Do we want to show our result for LSST vs SRD assumption of $0.95/D(z)$?}

An alternative use might be galaxy evolution studies, where these results provide a convenient cross-check formula. Our full parameter chains for the different limiting magnitudes are available upon request. \as{Is this the case?} \an{That would be totally fine with me.}
